Multi-dimensional Urban Network Percolation

### 2020-02-13 Docs/Papers/Percolation/submission/Raimbault2019/Raimbault2019.tex

%% Information provided by JIMIS:
%ISSN
% DOI 
\doi{10.18713/JIMIS-ddmmyy-v-a}{}
% reviewing process: first date = submission, second date = acceptance 
\review{Month-in-letters Day Year}{Month-in-letters Day Year}
% Volume and Year of the issue
\publication{N}{YYYY}

%% Information provided by the Editors:
% Issue Title 
\issue{Title of the interdisciplinary issue}
% List of k Editors
\editors{}

@@

\abstract{Network percolation has recently been proposed as a method to characterize the hierarchical structure of an urban system form the bottom-up. This paper proposes to extend urban network percolation in a multi-dimensional way, to take into account both urban form (spatial distribution of population) and urban functions (here as properties of transportation networks). The method is applied to the European urban system to reconstruct endogenous urban regions. The variable parametrization allows to consider patterns of optimization for two stylized contradictory sustainability indicators (economic performance and greenhouse gases emissions). This suggests a customizable spatial design of policies to develop sustainable territories.}

@@

Existing heuristics generally focus on a single dimension or property of the urban system. However, such systems are known to be highly multidimensional. For example, the morphological dimension of networks is complementary to the functional properties of the urban environment \citep{burger2012form}. The link between urban form and function remains in particular an open question \citep{batty1994fractal}. More generally, the inclusion of multiple dimension in urban analysis is still a research direction to be investigated, as in the case of agent-based models for example \citep{perez2016agent}. This paper addresses such a gap in the case of urban network percolation, by introducing a multi-dimensional percolation heuristic. The method allows combining different dimensions of the urban system, the same way that \cite{cottineau2018defining} link population density with commuting flows to produce multiple definitions of urban areas. We can indeed expect significantly different qualitative behaviors when switching from a single-dimension percolation to a multi-dimensional one, since the spatial structure of different urban dimensions are correlated but also largely complementary. If this is the case, the method then captures more aspects of the urban system and some of their interrelations.

@@

Case studies of targeted megacities have for example focused on the links between urban form and mobility or resources management \citep{sorensen2010megacities}. A significant literature has focused on the sustainability of mega urban regions with a qualitative approach, such as \cite{laquian2011planning} which establish a typology of these and corresponding governance structures in the Asian context. In the case of Europe, \cite{marull2013emerging} use econometric analysis to see the economic and ecological advantage of integrated urban regions. \cite{feng2018spatiotemporal} introduce a method to measure the level of polycentricity of an urban mega-region. \cite{su2017china} propose a framework to evaluate the performance of urban mega-regions, regarding economic, environmental, social and spatial dimensions.

@@

The rest of this paper is organized as follows: we first describe the multi-dimensional percolation heuristic, the data and variables to which it is applied and how the network is constructed, and the indicators used to characterized the sustainability of clusters produced. We then describe the results of applying this method to population and road network indicators for the whole European Union, focusing on the endogenous regions produced and their sustainability properties. We finally discuss possible developments and the implications of this methodology to the design of policies.

@@

%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[ht] 
%\resizebox{10cm}{7cm}
  {\includegraphics[width=\linewidth]{Fig1.png}}
  \centering
  \caption{Schematic representation of the multi-dimensional network percolation heuristic. Two layers are considered here, with similar nodes but different links in each. To each layer is associated a node variable. In each layer, links are created according to a distance threshold (percolation radius) and a variable value threshold. The final clusters (bottom in color) are obtained by considering the links present in each layer.\label{fig:method}}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%

@@

We do not call our method ``multi-layer percolation'', since nodes are common between layers. We show in Fig.~\ref{fig:method} for a schematic representation of the method. It can be implemented with a propagation heuristic or directly working on adjacency matrices. The rationale behind combining a thresholding for each layer variable with a distance thresholding relies on the idea that a first component for two points to interact is a geographical proximity. The second component is a strong enough intensity of the urban dimension captured by each layer, simultaneously for all layers considered. This recalls Tobler's first law of geography \citep{tobler2004first} in a multi-dimensional way.

@@

We use this data to construct a two layers abstract network: a layer which variable is given by population density, and a second layer which variable is given by a road network indicator. Nodes are the center of cells (thus disposed in space on a grid of step 5km). We test several possible networks by varying the road indicator taken into account for the second layer. In particular, we test the following indicators:
\begin{enumerate}
	\item number of edges $N_E$;
	\item number of vertices $N_V$;
	\item cyclomatic number $\mu$ which is defined by $\mu = N_E - N_V + c$ where $c$ is the number of connected components of the graph; this measure captures the number of distinct cycles in the graph
	\item euclidian efficiency $v_0$, defined by \cite{banos2012towards}, as the average rate between network distance and euclidian distance for all couples of links
\end{enumerate}

The choice of these measures is aimed at capturing basic aspects of network structure, and functional properties especially for the two last. Indeed, euclidian efficiency measures how the network is performant to link nodes, while the number of cycles is linked to redundancy of paths and in a way to robustness. These choice are arbitrary, but several aspects of transportation networks are still captured by these indicators. An increase of each is related to a more performant network regarding different dimensions, what is relevant for our application. A systematic study with other indicators such as centralities or accessibilities is out of the scope of this paper.

@@

Using these potential flows follows the logic of \cite{arbabi2019development} which shows a need for improved intra-city-region mobility in England and Wales. Considering the regions as entities in which such transportation development policies can more easily been developed, we look at the sustainability of different possible regions if these potential flows were realized. Varying the parameters $\gamma$ and $d_0$ allows to control for the economic activity considered (high $\gamma$ values correspond to high added-value activities) and the span of interactions through $d_0$.

@@

%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[ht] 
%\resizebox{10cm}{7cm}
  {\includegraphics[width=\linewidth]{Fig2.png}}
  \centering
  \caption{Percolation transition. On the left, we plot the size of the largest cluster in each configuration in terms of nodes, as a function of the percolation radius $r_0$. Color gives the other percolation parameters. On the right, the plot is similar but with the size relative to the size of the largest cluster obtained with the maximal radius in each configuration.\label{fig:percolation}}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[ht] 
%\resizebox{10cm}{7cm}
  {\includegraphics[width=\linewidth]{Fig3.png}}
  \centering
  \caption{Fractal dimension. We plot for each parametrization given by the curve color the evolution of the fractal dimension $\alpha$ as a function of $r_0$. Standard errors are not plotted for readability.\label{fig:fractaldim}}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%

@@

In its application to road networks by \cite{arcaute2016cities}, the structure of the national urban system for UK is captured by studying the percolation transition, i.e. the variation of the size of the largest cluster as a function of the percolation radius. As this signature is tightly linked to historical, cultural and geographical conditions, the application to different urban systems should yield different results. We study here this property, for different threshold parameter values. We make the radius vary between 8km and 100km with a one km step, have a fixed population threshold $\theta_P = 0.85$, test all road network indicators, and three network layer thresholds $\theta_N \in \{ 0 ; 0.8 ; 0.95 \}$. These values yield a good precision for the radius which is the most important variable to study transition and estimate fractal dimensions, while this population threshold is enough to provide a broad coverage for large radiuses (as shown by other explorations described below). Changing network variables and their threshold allows us to investigate how the transition behavior do change regarding the dimension considered and its intensity.

The absolute and relative sizes of the largest cluster are plotted in Fig.~\ref{fig:percolation} as a function of the percolation radius. This aspect first gives methodological information on multilayer percolation. Indeed, comparing the result with $\theta_N = 0$ (single layer percolation) with positive values of $\theta_N$ shows a significantly different behavior. As expected, absolute size are much smaller, but when looking at relative sizes we observe that the abrupt steps typical to percolation transitions have different distributions across the different parametrizations. This result is particularly interesting regarding the first axis of our research question, as it shows that the structure of clusters obtained is not only due to the population layer, and that the multi-dimensional percolation captures a complementary signal.

@@

where $N_c$ is the size of cluster $c$ and $\delta_c$ its diameter. These estimations are shown in Fig.~\ref{fig:fractaldim}. As a negative result, which could be due to the abstract nature of our network, a clear maximum in the value of the fractal dimension can not be found. Either it is located at a resolution that our method can not reach due to the minimal 5km limit imposed by the abstraction in the network construction, or it does not exist when coupling dimensions. Determining which assumption is more plausible is out of the scope of this paper.

@@

%%%%%%%%%%%%%%%%%%%%
\begin{figure}[h!] 
%\resizebox{10cm}{7cm}
{\includegraphics[width=\linewidth]{Fig4.png}}
  \centering
  \caption{Examples of obtained clusters for different parameter values. In the top-right case for example ($\theta_P = 0.9$, $\theta_N = 0.8$, variable \texttt{vcount},$r_0 = 8km$), we obtain the urban regions of West midlands and London in the UK, Randstad merged with Rhein-Rhur and Rhein-Main in Germany, Paris in France, also with capital cities such as Copenhaguen, Stockholm and Helsinki. There is no cluster in South Europe in that case, due to the high population density threshold.\label{fig:exclusters}}
\end{figure}
%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%
\begin{figure}[h!] 
%\resizebox{10cm}{7cm}
  {\includegraphics[width=0.7\linewidth]{Fig5.png}}
  \centering
  \caption{Point clouds of region-level indicators, namely population and emissions, for different parametrizations, given by the color. Each point represent an endogenous urban region.\label{fig:paretos}}
\end{figure}
%%%%%%%%%%%%%%%%%%%%

@@

We obtain very different endogenous morphologies for the different parametrizations. Maps reveal that some configurations resemble the actual distribution of European mega-city regions, which are functionally integrated polycentric urban areas \citep{hall2006polycentric}. These are here defined endogenously from the bottom-up and have a priori no reason to coincide with these functional regions. We show some examples in Fig.~\ref{fig:exclusters}. The first map of this figure, obtained for high population and network thresholds ($\theta_P = 0.95$ and $\theta_N = 0.9$), but a low radius $r_0 = 8$km and edge count $N_E$ to define the road network layer, include several mega-city regions described by \citep{hall2006polycentric}, namely London metropolitan area, the Randstad in Netherland, the Rhein-Main and Rhein-Ruhr in Germany, Greater Paris in France, Brussels area in Belgium. The same parameters with $\theta_N = 0$ yield not exactly the same regions, as confirmed by the transition curves in Fig.~\ref{fig:percolation}, what means that our approach taking into account two dimensions may capture effective processes of mega-city regions, in particular by including the road network which is crucial as these regions are integrated in terms of flows. The bottom-left map show an example of large clusters emerging in UK and in the center of Europe, the South remaining largely disconnected. Finally, the last map shows the result obtained with a very high radius $r_0 = 54$km, with a giant cluster spanning most of Europe. UK is still disconnected and the transition where it connects happens at $r_0 = 55$km. This does not necessarily mean that UK should be disconnected from continental Europe, as we considered geographic distances only, hiding the high speed connection of the Channel tunnel.

The behavior of sustainability indicators for different population, network and distance thresholds yield different distributions of performances across clusters within a configuration but also between configurations. Before considering the flow-based indicators described above, we can already study basic summary measures such as population $P_c$ and effective emissions $EM_c$. We show in Fig.~\ref{fig:paretos} point clouds of $\log EM_c$ against $1 - \log P_c$ for some configurations. Indeed, regarding the population it contains, an area can be more or less efficient in terms of emissions. Seeing the population as an objective to be maximized (thus the plotted value to be minimized) while the emissions must be minimized, we observe a Pareto front for all points (i.e. all clusters across all configurations). Given different dimensions to be minimized, a Pareto front consists of the points which are not Pareto-dominated by any other point, i.e. that there exists no other point performing best on all objectives. In practice, this yields optimization compromises in the context of multi-objective optimization, when no aggregation of the dimensions is possible or desirable. In Fig.~\ref{fig:paretos}, the front is the lower bound of the point cloud. We also find no dominating point for each configuration, i.e. that considering point clouds of a single color, a Pareto front with more than one point is still present. Some clusters are therefore optimal compromises in the Pareto sense in each configuration, while some are dominated and thus not efficient. This confirms that urban systems are generally compromises between multiple objectives.

@@

%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[!ht] 
%\resizebox{10cm}{7cm}
  {\includegraphics[width=\linewidth]{Fig7.png}}
  \centering  
  \caption{Relative potential emissions $\sum_c g_c$ against relative potential economic unefficiency $\sum_c e_c$ (both indicators should be minimized), for varying values of $\gamma_G$ (columns) and $d_G$ (rows). Color level gives the value of PC1, whereas point size gives the share of total population contained within considered areas.\label{fig:paretos-relative}}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%

@@

We show in Fig.~\ref{fig:emissions-pc1} the value of $\sum_c \tilde{g}_c$ as a function of the first morphological principal component, for extreme values of $\gamma$ and $d_0$. There seems to exist an optimal intermediate value for PC1 regarding the minimization of normalized indicator for emissions only. This would correspond to an intermediate level of monocentricity, meaning that urban areas which are too polycentric and spread would emit more, but also areas that are too much monocentric. This behavior does not occur for long-range $d_0 = 100$km and low-hierarchy $\gamma=0.5$ interactions. The mostly monocentric but emitting configurations capture most of population (given by the level of color), whereas the intermediate configuration capture around half of the population, what means that these low-emissions potential urban regions can cover a significant part of European population.

However, when considering both emissions and economic indicators, urban form then acts as a compromise variable. We show in Fig.~\ref{fig:paretos-relative} the point clouds of $\sum_c g_c$ against $\sum_c e_c$, which produce clear Pareto fronts, which shape varies with $\gamma$ and $d_0$. As the color level gives the value of PC1, we can see the points on the different fronts with very different morphological properties. In some cases, highly monocentric areas (yellow points) can be a good compromise, whereas the intermediate optimal for emissions shown before may yield highly inefficient areas (dominated green points). For example, considering the fronts for $\gamma = 2$ which have both very similar shape, the points with the lowest emissions are on the top-left of the front and correspond to the optimal unveiled in Fig.~\ref{fig:emissions-pc1}. These have however a very low economic efficiency (high inefficiency) and small improvements can be done with the points below, before switching to a totally different urban form with a high value of PC1 (yellow points, highly monocentric). Increasing more the economic efficiency is then at the price of much more emissions, with more polycentric areas. This analysis therefore unveils morphological trade-offs, confirming that there is no optimal urban form, but different compromises regarding the conflicting sustainability indicators.

@@

Further work may consist in the use of calibration heuristics to find in a more robust way optimal parameter values. The OpenMOLE model exploration platform provides a transparent access to genetic algorithms for multi-objective optimization \citep{reuillon2013openmole}. The use of such calibration algorithms would allow to unveil the effective form of Pareto fronts, that we may have missed here through the grid sampling.

@@

Our work suggests the possibility to design policies in terms of regional integration to increase the sustainability of mega-city regions. The way such results could actually be transferred to policy-making recommandations remains an open question, but Pareto-optimal configurations can be used for the planning of regional transportation networks for example, or to design policies for the distribution of subsidies. Indeed, privileging some infrastructure developments but also collaborations between urban centers can be seen as an aspect of a small scale planning, or territorial strategy. As we integrated potential flows that would result from such development, and consider their economic and emissions consequences, and did it in an endogenous way, we suggest that evidence-based strategies for territorial development at the European level could be inspired by this work. This would naturally imply a more thorough data integration, model calibration and operationalization.

@@

In conclusion, our multilayer percolation approach captures in a way the multi-dimensionality of urban systems and a link between form and function in urban system. In particular, in our application on the bilayer case of an abstract network constructed from population density and road network indicators, it is shown to capture a different structure than in the unidimensional case. Its application to the issue of sustainable mega-city regions show how its properties of constructing urban clusters from the bottom-up can be used to study sustainability issues. This work also illustrates the importance of following data-driven paradigms even when developing, as what is understood of the behavior of the heuristic is through its application to real data and issues.

@@

\begin{thebibliography}{}

\bibitem[\protect\citeauthoryear{Aleta and Moreno}{Aleta and
  Moreno}{2018}]{aleta2018multilayer}
Aleta A., Moreno Y. (2018).
\newblock Multilayer networks in a nutshell.
\newblock {\em Annual Review of Condensed Matter Physics\/}.

\bibitem[\protect\citeauthoryear{Arbabi, Mayfield, and McCann}{Arbabi
  \textit{et~al.}}{2019}]{arbabi2019development}
Arbabi H., Mayfield M., McCann P. (2019).
\newblock On the development logic of city-regions: inter-versus intra-city
  mobility in england and wales.
\newblock {\em Spatial Economic Analysis\/}, 1--20.

\bibitem[\protect\citeauthoryear{Arcaute, Molinero, Hatna, Murcio, Vargas-Ruiz,
  Masucci, and Batty}{Arcaute \textit{et~al.}}{2016}]{arcaute2016cities}
Arcaute E., Molinero C., Hatna E., Murcio R., Vargas-Ruiz C., Masucci A.~P.,
  Batty M. (2016).
\newblock Cities and regions in britain through hierarchical percolation.
\newblock {\em Royal Society open science\/}~{\em 3\/}(4), 150691.

\bibitem[\protect\citeauthoryear{Banos and Genre-Grandpierre}{Banos and
  Genre-Grandpierre}{2012}]{banos2012towards}
Banos A., Genre-Grandpierre C. (2012).
\newblock Towards new metrics for urban road networks: Some preliminary
  evidence from agent-based simulations.
\newblock In {\em Agent-based models of geographical systems}, pp.\  627--641.
  Springer.

\bibitem[\protect\citeauthoryear{Barth{\'e}lemy}{Barth{\'e}lemy}{2011}]{barthelemy2011spatial}
Barth{\'e}lemy M. (2011).
\newblock Spatial networks.
\newblock {\em Physics Reports\/}~{\em 499\/}(1-3), 1--101.

\bibitem[\protect\citeauthoryear{Batty and Longley}{Batty and
  Longley}{1994}]{batty1994fractal}
Batty M., Longley P.~A. (1994).
\newblock {\em Fractal cities: a geometry of form and function}.
\newblock Academic press.

\bibitem[\protect\citeauthoryear{Bettencourt, Lobo, Helbing, K{\"u}hnert, and
  West}{Bettencourt \textit{et~al.}}{2007}]{bettencourt2007growth}
Bettencourt L.~M., Lobo J., Helbing D., K{\"u}hnert C., West G.~B. (2007).
\newblock Growth, innovation, scaling, and the pace of life in cities.
\newblock {\em Proceedings of the national academy of sciences\/}~{\em
  104\/}(17), 7301--7306.

\bibitem[\protect\citeauthoryear{Boccaletti, Bianconi, Criado, Del~Genio,
  G{\'o}mez-Gardenes, Romance, Sendina-Nadal, Wang, and Zanin}{Boccaletti
  \textit{et~al.}}{2014}]{boccaletti2014structure}
Boccaletti S., Bianconi G., Criado R., Del~Genio C.~I., G{\'o}mez-Gardenes J.,
  Romance M., Sendina-Nadal I., Wang Z., Zanin M. (2014).
\newblock The structure and dynamics of multilayer networks.
\newblock {\em Physics Reports\/}~{\em 544\/}(1), 1--122.

\bibitem[\protect\citeauthoryear{Burger and Meijers}{Burger and
  Meijers}{2012}]{burger2012form}
Burger M., Meijers E. (2012).
\newblock Form follows function? linking morphological and functional
  polycentricity.
\newblock {\em Urban studies\/}~{\em 49\/}(5), 1127--1149.

>>>>

%% Information provided by JIMIS:
%ISSN
% DOI 
\doi{10.18713/JIMIS-180719-5-5}{}
% reviewing process: first date = submission, second date = acceptance 
\review{March 17, 2019}{October 23, 2019}
% Volume and Year of the issue
\publication{5}{2019}

%% Information provided by the Editors:
% Issue Title 
\issue{Graph and network analysis}
% List of k Editors
\editors{Vincent Labatut}

\pgfplotsset{compat=1.14}

@@

\abstract{Network percolation has recently been proposed as a method to characterize the hierarchical structure of an urban system from the bottom-up. This paper proposes to extend urban network percolation in a multi-dimensional way, to take into account both urban form (spatial distribution of population) and urban functions (here as properties of transportation networks). The method is applied to the European urban system to reconstruct endogenous urban regions. The variable parametrization allows to consider patterns of optimization for two stylized contradictory sustainability indicators (economic performance and greenhouse gases emissions). This suggests a customizable spatial design of policies to develop sustainable territories.}

@@

Existing heuristics generally focus on a single dimension or property of the urban system. However, such systems are known to be highly multidimensional. For example, the morphological dimension of networks is complementary to the functional properties of the urban environment \citep{burger2012form}. The link between urban form and function remains in particular an open question \citep{batty1994fractal}. More generally, the inclusion of multiple dimensions in urban analysis is still a research direction to be investigated, as in the case of agent-based models for example \citep{perez2016agent}. This paper addresses such a gap in the case of urban network percolation, by introducing a multi-dimensional percolation heuristic. The method allows combining different dimensions of the urban system, the same way that \cite{cottineau2018defining} link population density with commuting flows to produce multiple definitions of urban areas. We can indeed expect significantly different qualitative behaviors when switching from a single-dimension percolation to a multi-dimensional one, since the spatial structure of different urban dimensions are correlated but also largely complementary. If this is the case, the method then captures more aspects of the urban system and some of their interrelations.

@@

Case studies of targeted megacities have for example focused on the links between urban form and mobility or resources management \citep{sorensen2010megacities}. A significant literature has focused on the sustainability of mega urban regions with a qualitative approach, such as \cite{laquian2011planning} which establishes a typology of these and corresponding governance structures in the Asian context. In the case of Europe, \cite{marull2013emerging} use econometric analysis to see the economic and ecological advantage of integrated urban regions. \cite{feng2018spatiotemporal} introduce a method to measure the level of polycentricity of an urban mega-region. \cite{su2017china} propose a framework to evaluate the performance of urban mega-regions, regarding economic, environmental, social and spatial dimensions.

@@

The rest of this paper is organized as follows: we first describe the multi-dimensional percolation heuristic, the data and variables to which it is applied and how the network is constructed, and the indicators used to characterize the sustainability of clusters produced. We then describe the results of applying this method to population and road network indicators for the whole European Union, focusing on the endogenous regions produced and their sustainability properties. We finally discuss possible developments and the implications of this methodology to the design of policies.

@@

%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[!ht] 
%\resizebox{10cm}{7cm}
  {\includegraphics[width=0.75\linewidth]{Fig1.png}}
  \centering
  \caption{Schematic representation of the multi-dimensional network percolation heuristic. Two layers are considered here, with similar nodes but different links in each. To each layer is associated a node variable. In each layer, links are created according to a distance threshold (percolation radius) and a variable value threshold. The final clusters (bottom in color) are obtained by considering the links present in each layer.\label{fig:method}}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%

@@

We do not call our method ``multi-layer percolation'', since nodes are common between layers. We show in Figure~\ref{fig:method} for a schematic representation of the method. It can be implemented with a propagation heuristic or directly working on adjacency matrices. The rationale behind combining a thresholding for each layer variable with a distance thresholding relies on the idea that a first component for two points to interact is a geographical proximity. The second component is a strong enough intensity of the urban dimension captured by each layer, simultaneously for all layers considered. This recalls Tobler's first law of geography \citep{tobler2004first} in a multi-dimensional way.

@@

We use this data to construct a two layers abstract network: a layer which variable is given by population density, and a second layer which variable is given by a road network indicator. Nodes are the center of cells (thus disposed in space on a grid of step 5km). We test several possible networks by varying the road indicator taken into account for the second layer. In particular, we test the following indicators:
\begin{enumerate}
	\item number of edges $N_E$;
	\item number of vertices $N_V$;
	\item cyclomatic number $\mu$ which is defined by $\mu = N_E - N_V + c$ where $c$ is the number of connected components of the graph; this measure captures the number of distinct cycles in the graph;
	\item euclidian efficiency $v_0$, defined by \cite{banos2012towards}, as the average rate between network distance and euclidian distance for all couples of links.
\end{enumerate}

The choice of these measures is aimed at capturing basic aspects of network structure, and functional properties especially for the two last. Indeed, euclidian efficiency measures how the network is performant to link nodes, while the number of cycles is linked to redundancy of paths and in a way to robustness. These choices are arbitrary, but several aspects of transportation networks are still captured by these indicators. An increase of each is related to a more performant network regarding different dimensions, what is relevant for our application. A systematic study with other indicators such as centralities or accessibilities is out of the scope of this paper.

@@

Using these potential flows follows the logic of \cite{arbabi2019development} which show a need for improved intra-city-region mobility in England and Wales. Considering the regions as entities in which such transportation development policies can more easily been developed, we look at the sustainability of different possible regions if these potential flows were realized. Varying the parameters $\gamma$ and $d_0$ allows controlling for the economic activity considered (high $\gamma$ values correspond to high added-value activities) and the span of interactions through $d_0$.

@@

%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[!ht] 
%\resizebox{10cm}{7cm}
  {\includegraphics[width=\linewidth]{Fig2.png}}
  \centering
  \caption{Percolation transition. On the left, we plot the size of the largest cluster in each configuration in terms of nodes, as a function of the percolation radius $r_0$. Color gives the other percolation parameters. On the right, the plot is similar but with the size relative to the size of the largest cluster obtained with the maximal radius in each configuration.\label{fig:percolation}}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[!ht] 
%\resizebox{10cm}{7cm}
  {\includegraphics[width=\linewidth]{Fig3.png}}
  \centering
  \caption{Fractal dimension. We plot for each parametrization given by the curve color the evolution of the fractal dimension $\alpha$ as a function of $r_0$. Standard errors are not plotted for readability.\label{fig:fractaldim}}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%

@@

In its application to road networks by \cite{arcaute2016cities}, the structure of the national urban system for UK is captured by studying the percolation transition, i.e. the variation of the size of the largest cluster as a function of the percolation radius. As this signature is tightly linked to historical, cultural and geographical conditions, the application to different urban systems should yield different results. We study here this property, for different threshold parameter values. We make the radius vary between 8km and 100km with a one km step, have a fixed population threshold $\theta_P = 0.85$, test all road network indicators, and three network layer thresholds $\theta_N \in \{ 0 ; 0.8 ; 0.95 \}$. These values yield a good precision for the radius which is the most important variable to study transition and estimate fractal dimensions, while this population threshold is enough to provide a broad coverage for large radiuses (as shown by other explorations described below). Changing network variables and their threshold allows us to investigate how the transition behavior does change regarding the dimension considered and its intensity.

The absolute and relative sizes of the largest cluster are plotted in Figure~\ref{fig:percolation} as a function of the percolation radius. This aspect first gives methodological information on multilayer percolation. Indeed, comparing the result with $\theta_N = 0$ (single layer percolation) with positive values of $\theta_N$ shows a significantly different behavior. As expected, absolute size are much smaller, but when looking at relative sizes we observe that the abrupt steps typical to percolation transitions have different distributions across the different parametrizations. This result is particularly interesting regarding the first axis of our research question, as it shows that the structure of clusters obtained is not only due to the population layer, and that the multi-dimensional percolation captures a complementary signal.

@@

where $N_c$ is the size of cluster $c$ and $\delta_c$ its diameter. These estimations are shown in Figure~\ref{fig:fractaldim}. As a negative result, which could be due to the abstract nature of our network, a clear maximum in the value of the fractal dimension can not be found. Either it is located at a resolution that our method can not reach due to the minimal 5km limit imposed by the abstraction in the network construction, or it does not exist when coupling dimensions. Determining which assumption is more plausible is out of the scope of this paper.

@@

%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!ht] 
%\resizebox{10cm}{7cm}
{\includegraphics[width=\linewidth]{Fig4.png}}
  \centering
  \caption{Examples of obtained clusters for different parameter values. In the top-right case for example ($\theta_P = 0.9$, $\theta_N = 0.8$, variable \texttt{vcount},$r_0 = 8km$), we obtain the urban regions of West midlands and London in the UK, Randstad merged with Rhein-Rhur and Rhein-Main in Germany, Paris in France, also with capital cities such as Copenhaguen, Stockholm and Helsinki. There is no cluster in South Europe in that case, due to the high population density threshold.\label{fig:exclusters}}
\end{figure}
%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!ht] 
%\resizebox{10cm}{7cm}
  {\includegraphics[width=0.7\linewidth]{Fig5.png}}
  \centering
  \caption{Point clouds of region-level indicators, namely population and emissions, for different parametrizations, given by the color. Each point represent an endogenous urban region.\label{fig:paretos}}
\end{figure}
%%%%%%%%%%%%%%%%%%%%

@@

We obtain very different endogenous morphologies for the different parametrizations. Maps reveal that some configurations resemble the actual distribution of European mega-city regions, which are functionally integrated polycentric urban areas \citep{hall2006polycentric}. These are here defined endogenously from the bottom-up and have a priori no reason to coincide with these functional regions. We show some examples in Figure~\ref{fig:exclusters}. The first map of this figure, obtained for high population and network thresholds ($\theta_P = 0.95$ and $\theta_N = 0.9$), but a low radius $r_0 = 8$km and edge count $N_E$ to define the road network layer, include several mega-city regions described by \cite{hall2006polycentric}, namely London metropolitan area, the Randstad in Netherland, the Rhein-Main and Rhein-Ruhr in Germany, Greater Paris in France, Brussels area in Belgium. The same parameters with $\theta_N = 0$ yield not exactly the same regions, as confirmed by the transition curves in Figure~\ref{fig:percolation}, what means that our approach taking into account two dimensions may capture effective processes of mega-city regions, in particular by including the road network which is crucial as these regions are integrated in terms of flows. The bottom-left map show an example of large clusters emerging in UK and in the center of Europe, the South remaining largely disconnected. Finally, the last map shows the result obtained with a very high radius $r_0 = 54$km, with a giant cluster spanning most of Europe. UK is still disconnected and the transition where it connects happens at $r_0 = 55$km. This does not necessarily mean that UK should be disconnected from continental Europe, as we considered geographic distances only, hiding the high speed connection of the Channel tunnel.

The behavior of sustainability indicators for different population, network and distance thresholds yield different distributions of performances across clusters within a configuration but also between configurations. Before considering the flow-based indicators described above, we can already study basic summary measures such as population $P_c$ and effective emissions $EM_c$. We show in Figure~\ref{fig:paretos} point clouds of $\log EM_c$ against $1 - \log P_c$ for some configurations. Indeed, regarding the population it contains, an area can be more or less efficient in terms of emissions. Seeing the population as an objective to be maximized (thus the plotted value to be minimized) while the emissions must be minimized, we observe a Pareto front for all points (i.e. all clusters across all configurations). Given different dimensions to be minimized, a Pareto front consists of the points which are not Pareto-dominated by any other point, i.e. that there exists no other point performing best on all objectives. In practice, this yields optimization compromises in the context of multi-objective optimization, when no aggregation of the dimensions is possible or desirable. In Figure~\ref{fig:paretos}, the front is the lower bound of the point cloud. We also find no dominating point for each configuration, i.e. that considering point clouds of a single color, a Pareto front with more than one point is still present. Some clusters are therefore optimal compromises in the Pareto sense in each configuration, while some are dominated and thus not efficient. This confirms that urban systems are generally compromises between multiple objectives.

@@

%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[!ht] 
%\resizebox{10cm}{7cm}
  {\includegraphics[width=\linewidth]{Fig7.png}}
  \centering  
  \caption{Relative potential emissions $\sum_c g_c$ against relative potential economic inefficiency $\sum_c e_c$ (both indicators should be minimized), for varying values of $\gamma_G$ (columns) and $d_G$ (rows). Color level gives the value of PC1, whereas point size gives the share of total population contained within considered areas.\label{fig:paretos-relative}}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%

@@

We show in Figure~\ref{fig:emissions-pc1} the value of $\sum_c \tilde{g}_c$ as a function of the first morphological principal component, for extreme values of $\gamma$ and $d_0$. There seems to exist an optimal intermediate value for PC1 regarding the minimization of the normalized indicator for emissions only. This would correspond to an intermediate level of monocentricity, meaning that urban areas which are too polycentric and spread would emit more, but also areas that are too much monocentric. This behavior does not occur for long-range $d_0 = 100$km and low-hierarchy $\gamma=0.5$ interactions. The mostly monocentric but emitting configurations capture most of population (given by the level of color), whereas the intermediate configurations capture around half of the population, what means that these low-emissions potential urban regions can cover a significant part of European population.

However, when considering both emissions and economic indicators, urban form then acts as a compromise variable. We show in Figure~\ref{fig:paretos-relative} the point clouds of $\sum_c g_c$ against $\sum_c e_c$, which produce clear Pareto fronts, which shape varies with $\gamma$ and $d_0$. As the color level gives the value of PC1, we can see the points on the different fronts with very different morphological properties. In some cases, highly monocentric areas (yellow points) can be a good compromise, whereas the intermediate optimal for emissions shown before may yield highly inefficient areas (dominated green points). For example, considering the fronts for $\gamma = 2$ which have both very similar shape, the points with the lowest emissions are on the top-left of the front and correspond to the optimal unveiled in Figure~\ref{fig:emissions-pc1}. These have however a very low economic efficiency (high inefficiency) and small improvements can be done with the points below, before switching to a totally different urban form with a high value of PC1 (yellow points, highly monocentric). Increasing more the economic efficiency is then at the price of much more emissions, with more polycentric areas. This analysis therefore unveils morphological trade-offs, confirming that there is no optimal urban form, but different compromises regarding the conflicting sustainability indicators.

@@

Further work may consist in the use of calibration heuristics to find in a more robust way optimal parameter values. The OpenMOLE model exploration platform provides a transparent access to genetic algorithms for multi-objective optimization \citep{reuillon2013openmole}. The use of such calibration algorithms would allow unveiling the effective form of Pareto fronts, that we may have missed here through the grid sampling.

@@

Our work suggests the possibility to design policies in terms of regional integration to increase the sustainability of mega-city regions. The way such results could actually be transferred to policy-making recommandations remains an open question, but Pareto-optimal configurations can be used for the planning of regional transportation networks for example, or to design policies for the distribution of subsidies. Indeed, privileging some infrastructure developments but also collaborations between urban centers can be seen as an aspect of a small scale planning, or territorial strategy. As we integrated potential flows that would result from such development, and considered their economic and emissions consequences, and did it in an endogenous way, we suggest that evidence-based strategies for territorial development at the European level could be inspired by this work. This would naturally imply a more thorough data integration, model calibration and operationalization.

@@

In conclusion, our multilayer percolation approach captures in a way the multi-dimensionality of urban systems and a link between form and function in urban system. In particular, in our application on the bilayer case of an abstract network constructed from population density and road network indicators, it is shown to capture a different structure than in the unidimensional case. Its application to the issue of sustainable mega-city regions show how its properties of constructing urban clusters from the bottom-up can be used to study sustainability issues. This work also illustrates the importance of following data-driven paradigms even when developing new methods, as what is understood of the behavior of the heuristic is through its application to real data and issues.

@@

\bibliographystyle{jimis-en}
\bibliography{Raimbault2019}
